---
title: "Final Project Proposal - DATA 608"
author: "John Mazon"
date: "10/24/2021"
output: html_document
---

## DATA 608 Final Project Proposal


## Project Background
I will be utilizing the the NYPD Shooting Incident Data dataset which includes list of every shooting incident that occurred in NYC going back to 2006 through the end of the previous calendar year.

This is a breakdown of every shooting incident that occurred in NYC going back to 2006 through the end of the previous calendar year. This data is manually extracted every quarter and reviewed by the Office of Management Analysis and Planning before being posted on the NYPD website. Each record represents a shooting incident in NYC and includes information about the event, the location and time of occurrence. In addition, information related to suspect and victim demographics is also included. This data can be used by the public to explore the nature of shooting/criminal activity. 

Goal of the project will be to see what factors are involved and which trends are seen in gun violence for NYC
We are provided with many data points such as location, race, gender and date of occurrence. 

## Data Sources

https://data.cityofnewyork.us/Public-Safety/NYPD-Shooting-Incident-Data-Historic-/833y-fsy8

```{r}
df <- read.csv("https://raw.githubusercontent.com/johnm1990/DATA608/main/NYPD_Shooting_Incident_Data__Historic_.csv")


```

The datasource is from DATA.CITYOFNEWYORK, some parameters include borough, timeframe(back to 2006 through the end of the previous calendar year). Data shows many helpful insights on discovering which boroughs are most affected and how frequent this violence occurrs. Please view a "head" section of the data

```{r}
head(df)

````

## Supplementary Data

There exist structures in city government to combat this kind of violence such as NYC Office to prevent gun violence. Although gun violence has decreased to historically low levels in recent years, it remains a pervasive issue in small pockets of neighborhoods around New York City. The data shows that a limited number of communities have accounted for roughly 50% of NYC’s shooting incidents.  

In 2014, the NYC Council and the de Blasio administration launched the NYC Crisis Management System (CMS). This initiative simultaneously operates across 22 communities citywide in neighborhoods significantly impacted by gun violence and provides violence intervention and support systems. CMS consists of more than 50 community-based organizations providing various services to support those at the highest risk. 

Research has shown that when individuals closest to the issues are an integral part of the development and implementation of solutions, legitimacy and sustained norm change is possible. Through the de Blasio administration’s investment in the data-driven, community-based strategies, these facts continue to be affirmed through the Office to Prevent Gun Violence (OPGV).

## Analysis and Visualization

A shiny app will be utilized to view the distribution of gun violence in nyc. This will help to visualize any patterns. 


## Possibly useful libraries

    knitr
    plyr
    dplyr
    sqldf
    data.table
    DT
    kableExtra
    ggplot2
    reshape2
    ggplot2
    plotly
    graphics
   
   

